---
title: "We've all been foxed by Leicester"
output:
  html_document:
    theme: sandstone
    highlight: pygments
    include:
      after_body: footer.html
---

The season has been bonkers -- we all know that. But, by way of confirmation, this is the first PremPredict league where everyone's worst team is the same: Leicester City. 

Congratulations therefore go to the tinker man, Kanté, Mahrez and Vardy. But they should also go to Roger Gathercole, who leads the way in our competition. Somehow, Beth and I are currently also in the money.

Here are the standings:

```{r thecode, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, error=TRUE}

setwd("~/GitHub/p0bs.github.io/")

library(readr)
library(dplyr)
library(stringr)
library(tidyr)
library(knitr)
library(htmlwidgets)
library(DT)
library(magrittr)

# Start by running this dimmed code (as PP-thin.R):
# library(rvest)
# library(xml2)
# leagueTable <- read_html("http://www.theguardian.com/football/premierleague/table")
# clubOrder <- leagueTable %>% 
#   html_nodes(".table--striped") %>% 
#   .[[1]] %>% 
#   html_table()
# saveRDS(clubOrder, "clubOrder.rds")  # Add date
# Then move clubOrder.rds to appropriate GitHub drive

clubOrder <- readRDS("clubOrder.rds")  # Add date

# Rename some teams to make the league table consistent with our inputs
clubOrder$Team[clubOrder$Team=="Leicester"] <- "Leicester City"
clubOrder$Team[clubOrder$Team=="Spurs"] <- "Tottenham Hotspur"
clubOrder$Team[clubOrder$Team=="AFC Bournemouth"] <- "Bournemouth"
clubOrder$Team[clubOrder$Team=="C Palace"] <- "Crystal Palace"
clubOrder$Team[clubOrder$Team=="Man City"] <- "Manchester City"
clubOrder$Team[clubOrder$Team=="Man Utd"] <- "Manchester United"
clubOrder$Team[clubOrder$Team=="West Ham"] <- "West Ham United"
clubOrder$Team[clubOrder$Team=="Stoke"] <- "Stoke City"
clubOrder$Team[clubOrder$Team=="Swansea"] <- "Swansea City"
clubOrder$Team[clubOrder$Team=="West Brom"] <- "West Bromwich Albion"
clubOrder$Team[clubOrder$Team=="Norwich"] <- "Norwich City"
clubOrder$Team[clubOrder$Team=="Newcastle"] <- "Newcastle United"

clubsABC <- sort(clubOrder$Team)

dataInput <- read_csv("PP-2015.csv")
clubStandings <- match(dataInput$Club, clubOrder$Team, 0)
topClub <- match(1, clubStandings, 0)

predictions <- dataInput[,-1]

bonus <- -50*(predictions[topClub,]==1)
ssq <- function(x){(x-clubStandings)^2}
squares <- apply(predictions,2,ssq)
score <- colSums(squares) + bonus
names <- colnames(score)
names1 <- str_replace_all(names, "_", " ")
worst <- apply(squares,2,max)
findWorst <- function(y){match(worst[y],squares[,y],0)}
worstClubNo <- sapply(1:ncol(predictions),findWorst)
worstClub <- clubsABC[worstClubNo]

output <- rbind(score, bonus, worst, worstClubNo)
output1 <- data.frame(names1,t(output))
row.names(output1) <- NULL 
output2 <- output1 %>% 
  mutate(worstClub=clubsABC[worstClubNo]) %>%
  select(-worstClubNo) 
colnames(output2) <- c("Names", "Scores", "Bonus", "WorstCost", "WorstClub")

report <- output2 %>% 
  arrange(Scores) %>% 
  select(Names, Scores, Bonus, WorstClub, WorstCost)
datatable(report, 
          rownames = FALSE, 
          options = list(dom = 't',
                         pageLength = 23,
                         order = list(list(1, 'asc'), list(0, 'asc'))))
```

Remember: if you ever want the latest standings, just check [here](https://robin.shinyapps.io/PremPredict).

I'll be in touch again in a few weeks. Until then, all the best ...

-- Robin

---

PS Here's my code to prove that all is kosher:

```{r thecode, eval=FALSE, echo=TRUE}
```

And, for completeness, here are your team-by-team predictions and each team's latest position in the league:

```{r thepredictions, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, error=TRUE}

Position <- as.vector(1:20)
clubsOrder <- clubOrder$Team
clubsOrder1 <- as.data.frame(cbind(clubsOrder, Position)) %>% 
  rename(Club = clubsOrder) %>% 
  left_join(dataInput, by = "Club")

datatable(clubsOrder1,
          rownames = FALSE,
          options = list(dom = 't',
                         pageLength = 20,
                         order = list(0, 'asc'),
                         columnDefs = list(list(
            className = 'dt-right', targets = 1))))
```

---
